---
title: "Tutorial: Calculate the Community Design poverty indicator"
author: "Abigail Stamm"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial: poverty indicator}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The purpose of this tutorial is to set up poverty data to create the NCDM traffic safety indicator, including: percent household poverty by census tract. It produces a data frame at the end.

To start, you will need to download your desired American Community Survey (ACS) data from the [US Census](https://www2.census.gov/geo/tiger/TIGER_DP/). Datasets on the site are organized by year. Click on each year of interest, then on the TRACT file with your state's FIPS code in its name. This will download the geodatabase. The code is designed to process only one geodatabase at a time. Other ACS geographies may also work, but they have not been tested. 

First, load the community design package.

```{r package, eval=FALSE}
library(cdccommdes)
# to learn more about the package, type 
citation("cdccommdes")        # suggested citation
browseVignettes("cdccommdes") # tutorials and documentation
?cdccommdes                   # help files and about page
```

Next, define your settings, including file path, geodatabase name, and layer name (the table inside the geodatabase).

```{r settings, eval=FALSE}
# my file path
# note the forward slashes (Windows users will need to change them manually)
my_path <- "c:/my_projects/acs"
# my state's census tract geodatabase, without the extensions
my_gdb <- "ACS_2018_5YR_TRACT_36_NEW_YORK"
my_layer <- "X17_POVERTY"    # table in the geodatabase
```

Last, run the function that will create the indicator. Everything is bundled in the same function to make it easy to use. The object produced, `f`, is a data.farme, or table, containing variables for number of households overall and in poverty and a percentage calculation. 

```{r processing, eval=FALSE}
# enter your settings from above
f <- process_poverty(readpath = my_path, gdb_name = my_gdb, 
                     layer_name = my_layer)
```

A generic save function is included in the package if you want to use it. It writes an XML file and an Rda (R data) object.

```{r saving, eval=FALSE}
# at this time, the XML file does not incorporate CDC metadata
save_generic(filename = "tract_safety_indicators", dataset = f, 
             savepath = my_path)
```


